---
title: "Gaia DR3, NGC Terzan 9"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

Use the [Gaia DR3](https://www.cosmos.esa.int/web/gaia-users/archive/programmatic-access) with R ~ 2 arcmin.

* Analyze the cluster stars and find the average distance in parsec, convert this to DM = m-M (use the formula). 
* Find the [M/H] value, which should be similar to [Fe/H] in the table.  
* Define a polygon search for each cluster and define a box search 1 deg x 1 deg around the cluster. 
* Can you identify any cluster stars (by parallax, proper motion, and [M/H]) which are several cluster radii away from the cluster center?

###### Variables used: `source_id`, `ra`, `dec`, `parallax`, `pmra`, `pmdec`, `bp_rp`, `mh_gspphot`, `radial_velocity`, `phot_g_mean_mag`, `pmra_error`,`pmdec_error`, `parallax_error`, `phot_g_mean_flux_over_error`, `phot_bp_mean_flux_over_error`, `phot_rp_mean_flux_over_error`.



```{r}
terzan_9 <- read.csv("1657871725305O-result.csv")
```

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(ggpubr)
```

#### Summary Statistics 

```{r}
summary(terzan_9)
```

#### Model 1

a. Use `lm()` to regress ra on dec and save the regression as `model_1`.

```{r}
model_1 <- lm(ra ~ dec, data = terzan_9)
```

b. Regression results from the first model using `summary()`.

An increase of one unit of dec is associated with an additional -0.46067 unit decrease in ra. This relationship is statistically significant at < 0.001.

```{r}
summary(model_1)
```

c. Plot results from `model_1`.

```{r}
ggplot(data = model_1, aes(x = dec, y = ra)) +
  geom_point(alpha=0.5, size=2, color = 'orange') +
  labs(y="ra", x="dec") +
  stat_smooth()
```

#### Graph 

```{r}
terzan_9 %>%
  ggplot(aes(dec,ra)) +
  geom_point(alpha=0.5, size=2, color = 'blue') +
  labs(y="ra", x="dec")
```

#### Model 2: (Proper motion and error)

a. Use `lm()` to regress pmra on pmdec and save the regression as `model_2` then pmra_error on pmdec_error and save as `pm_error`.

```{r}
model_2 <- lm(pmra ~ pmdec, data = terzan_9)
pm_error <- lm(pmra_error ~ pmdec_error, data = terzan_9)
```

b. Regression results from the second and proper motion error model using `summary()`.

`Model 2`: An increase of one unit of pmdec is associated with an additional  0.3932 unit increase in pmra. This relationship is statistically significant at < 0.01.

`pm_error`: An increase of one unit of pmdec error is associated with an additional 1.4531332 unit increase in pmra error. This relationship is statistically significant at < 0.001.


```{r}
summary(model_2)
summary(pm_error)
```

c. Plot results from `model_2` and `pm_error`.

```{r}
ggplot(data = model_2, aes(x = pmdec, y = pmra)) +
  geom_point(alpha=0.5, size=2, color = 'orange') +
  labs(y="pmra", x="pmdec") +
  stat_smooth()

ggplot(data = pm_error, aes(x = pmdec_error, y = pmra_error)) +
  geom_point(alpha=0.5, size=2, color = 'orange') +
  labs(y="pmra", x="pmdec") +
  stat_smooth()
```

#### Graph 

```{r}
terzan_9 %>%
  ggplot(aes(pmdec,pmra)) +
  geom_point(alpha=0.5, size=2, color = 'blue') +
  labs(y="pmra", x="pmdec") 
```

#### Model 3: (G-mean and bp/rp error)

a. Use `lm()` to regress phot_g_mean_mag on bp_rp and save the regression as `model_3` then phot_g_mean_flux_over_error on phot_bp_mean_flux_over_error and phot_rp_mean_flux_over_error then save it as `phot_error`.

```{r}
model_3 <- lm(phot_g_mean_mag ~ bp_rp, data = terzan_9)
phot_error <- lm(phot_g_mean_flux_over_error ~ phot_bp_mean_flux_over_error + phot_rp_mean_flux_over_error, data = terzan_9)
```

b. Regression results from the third model using `summary()`.

An increase of one unit of bp_rp is associated with an additional -0.57164 unit decrease in phot_g_mean_mag. This relationship is statistically significant at < 0.001.

An increase of one unit of bp/rp error is associated with an additional 2.53272 unit and 1.36537 increase in phot_g_mean_mag error. This relationship is both statistically significant at < 0.001.


```{r}
summary(model_3)
summary(phot_error)
```

c. Plot results from `model_3` and `phot_error`.

```{r}
ggplot(data = model_3, aes(x = bp_rp, y = phot_g_mean_mag)) +
  geom_point(alpha=0.5, size=2, color = 'orange') +
  labs(y="phot_g_mean_mag", x="bp_rp") +
  stat_smooth()

ggplot(data = phot_error, aes(x = phot_bp_mean_flux_over_error + phot_rp_mean_flux_over_error, y = phot_g_mean_flux_over_error)) +
  geom_point(alpha=0.5, size=2, color = 'orange') +
  labs(y="phot_g_mean_mag_error", x="bp_rp_error") +
  stat_smooth()
```

#### Graph 

```{r}
terzan_9 %>%
  ggplot(aes(bp_rp,phot_g_mean_mag)) +
  geom_point(alpha=0.5, size=2, color = 'blue') +
  labs(y="phot_g_mean_mag", x="bp_rp")
```

#### Metallicity Over Hydrogen

```{r}
ggplot(terzan_9, aes(mh_gspphot)) +
  geom_histogram(bins = 30)
```


#### Radio Velocity

```{r}
ggplot(terzan_9, aes(radial_velocity)) +
  geom_histogram()
```

#### Polygon search for Terzan 9.

(Switch over to Jupyter Lab.)

```{r}
ps_terzan_9 <- terzan_9 
select(ps_ngc_6558, source_id, ra, dec, pmra, pmdec, parallax)
```









